Sequential algorithms to split and merge ultra-high resolution 3D images

### 2017-08-20 paper.tex
Minor fixes

The number of seeks is reported in Figure~\ref{fig:number-of-seeks},
for all algorithms and the corresponding models
(Equations~\ref{eq:naive-blocks}
to~\ref{eq:seeks-multiple-reads}). Note the logarithmic y scale. Error
bars are not reported as numbers of seeks were constant across all
repetitions. The average relative model errors are 12.7\% (Naive
blocks), 0\% (Naive slices), 3.3\% (Clustered reads), 26.8\% (Multiple
reads) and 0.9\% (Buffered slices), presumably explained by the fact
that the model assumes cubic blocks while we used non-cubic ones in
the experiment. For Multiple reads, our complexity analysis also
assumed that one of the 5 Cases in
Figure~\ref{fig:multiple-reads-cases} was used while they are usually
blended in practice. Overall, the model correctly explains the
observations.
\begin{figure}[h]
  \centering
  \includegraphics[width=\columnwidth]{figures/number-of-seeks.pdf}
  \hfill
  \caption{Number of seeks for all algorithms. y scale is logarithmic. Each algorithm is
    represented with a different color. Dark color is experimental
    value; bright color is model.}
  % Gnuplot file: scripts/number_of_seeks.gnuplot
\label{fig:number-of-seeks}
\end{figure}

As expected, the difference between Naive blocks and Naive slices is
tremendous, in the order of 50 million seeks. Regarding the difference between
Clustered reads and Multiple reads, it is consistent with our analysis. For
3~GB, the number of seeks in Clustered reads is 4 orders of magnitude
higher than for Multiple reads, and 5 orders of magnitude higher than
Naive slices. This huge difference comes from the fact that at 3~GB
Clustered reads are in Case 1. For 6~GB, 9~GB and 12~GB, Clustered
reads are in Case 2 and the difference with Multiple reads and
Buffered slices reduces. At 16~GB, all algorithms perform the same.

@@

% Slices and blocks
Figure~\ref{fig:merge-time} shows the merge time for all algorithms by
memory values. Naive blocks are 9.5 times slower than Naive slices on
HDD (6.7 times on SSD), which quantifies the effect of the targeted
problem. It also shows that SSDs might not be the only solution to
reduce the impact of seek time. In the remainder we use Naive blocks
and Naive slices as references to evaluate our algorithms.

@@

The total split time by algorithm is shown in
Figure~\ref{fig:split-time}. The difference between Naive blocks and
Naive slices is still significant (average ratio is 1.4 on SSD, 2.0 on
HDD) although less than for merging. On SSD, Clustered writes and
Multiple writes both perform similarly to Naive slices. On HDD,
Clustered writes are slightly slower than Multiple writes until
16~GB. Buffered slices don't provide any significant speed-up compared
to Naive slices.

@@

% File formats
Our results demonstrate that simple imaging
formats may be split and merged without performance loss compared to
more complex formats that can preserve spatial locality on disk,
for instance MINC 2.0 or the format based on space-filling curves
mentioned in~\cite{burns2013open}. This is of major interest in the
current open-science context since simpler formats favor data-sharing
and interoperability.  Moreover, our algorithms could potentially be
adapted to any split geometry, even though we demonstrated them on
slices and blocks only, while file formats inevitably assume a
particular geometry. For instance, the format in~\cite{burns2013open}
is not designed to naively split slices. However, we aimed to extract
all the blocks, whereas, ~\cite{burns2013open} aimed to extract a
single block from a large image. MINC 2.0 might also help with
on-the-fly compression for Clustered reads.

@@

% Parallel algorithms
High-resolution images are likely to be processed on computing
clusters, for instance using software from the Hadoop project, in
particular the Hadoop Distributed File
System~\cite{shvachko2010hadoop}. In this context parallel
split-and-merge algorithms would be beneficial, since the various
blocks of a large image could be uploaded to different disks
concurrently. In the same vein, ``re-spliting'' algorithms would be
beneficial in case an image already split needs to be split in a
different geometry. Designing such algorithms is part of our future
work. Clustered reads and multiple reads could be used as starting
points.

>>>>

The number of seeks is reported in Figure~\ref{fig:number-of-seeks},
for all algorithms and the corresponding models
(Equations~\ref{eq:naive-blocks}
to~\ref{eq:seeks-multiple-reads}). Note the logarithmic y scale. Error
bars are not reported as numbers of seeks were constant across all
repetitions. The average relative model errors are 12.7\% (Naive
blocks), 0\% (Naive slices), 3.3\% (Clustered reads), 26.8\% (Multiple
reads) and 0.9\% (Buffered slices), explained by the fact
that the model assumes cuboid blocks while we used non-cuboid ones in
the experiment. For Multiple reads, our complexity analysis also
assumed that one of the 5 Cases in
Figure~\ref{fig:multiple-reads-cases} was used while they are 
blended in practice. Overall, the model correctly explains the
observations.
\begin{figure}[h]
  \centering
  \includegraphics[width=\columnwidth]{figures/number-of-seeks.pdf}
  \hfill
  \caption{Number of seeks for all algorithms. y scale is logarithmic. Each algorithm is
    represented with a different color. Dark color is experimental
    value; bright color is model.}
  % Gnuplot file: scripts/number_of_seeks.gnuplot
\label{fig:number-of-seeks}
\end{figure}

As expected, the difference between Naive blocks and Naive slices is
tremendous, in the order of 50 million seeks. The difference between
Clustered reads and Multiple reads is consistent with our analysis. For
3~GB, the number of seeks in Clustered reads is 4 orders of magnitude
higher than for Multiple reads, and 5 orders of magnitude higher than
Naive slices. This huge difference comes from the fact that at 3~GB
Clustered reads are in Case 1. For 6~GB, 9~GB and 12~GB, Clustered
reads are in Case 2 and the difference with Multiple reads and
Buffered slices reduces. At 16~GB, all algorithms perform the same.

@@

% Slices and blocks
Figure~\ref{fig:merge-time} shows the merge time for all algorithms by
memory values. Naive blocks are 9.5 times slower than Naive slices on
HDD (6.7 times on SSD), which quantifies the effect of the targeted
problem.  In the remainder we use Naive blocks
and Naive slices as references to evaluate our algorithms.

@@

The total split time by algorithm is shown in
Figure~\ref{fig:split-time}. The difference between Naive blocks and
Naive slices is still significant (average ratio is 1.4 on SSD, 2.0 on
HDD) although less than for merging. On SSD, Clustered writes and
Multiple writes both perform similarly to Naive slices. On HDD,
Clustered writes are slightly slower than Multiple writes until
16~GB. Buffered slices provide no speed-up compared
to Naive slices.

@@

% File formats
Our results demonstrate that large images stored in simple 
formats may be split and merged without performance loss compared to
more complex formats that can preserve spatial locality on disk,
for instance MINC 2.0 or the format based on space-filling curves
mentioned in~\cite{burns2013open}. This is of major interest in the
current open-science context since simpler formats favor data-sharing
and interoperability.  Moreover, our algorithms could potentially be
adapted to any split geometry, even though we demonstrated them on
slices and blocks only, while file formats inevitably assume a
particular geometry. For instance, the format in~\cite{burns2013open}
is not designed to naively split slices. However, we aimed to extract
all the blocks, whereas, ~\cite{burns2013open} aimed to extract a
single block from a large image. MINC 2.0 might also help with
on-the-fly compression for Clustered reads.

@@

% Parallel algorithms
High-resolution images are likely to be processed on computing
clusters, for instance using software from the Hadoop project, in
particular the Hadoop Distributed File
System~\cite{shvachko2010hadoop}. In this context parallel
split-and-merge algorithms would be beneficial, since the various
blocks of a large image could be uploaded to different disks
concurrently. In the same vein, ``re-spliting'' algorithms would be
beneficial in case an image already split needs to be split in a
different geometry. Designing such algorithms is part of our future
work, in which Clustered reads and Multiple reads will be used as
starting points.
